Analyzing Students Use Metacognitive Strategies When Interacting with GenAI-Powered Chatbots in Programming Courses.

### 2025-01-30 ETHE/sn-article.tex
update paper and Java Labs

Recent advances in Generative Artificial Intelligence (GenAI) have created
new opportunities in both professional programming and education \citep{Puryear22}.
In programming courses, students can employ GenAI tools to improve their
understanding, receive personalized feedback, and access detailed explanations.
For example, GitHub Copilot, a tool that is integrated into development
environments, assists in providing real-time suggestions and accelerating code
writing, which could help streamline the learning process. Educational chatbots,
on the other hand, try to guide students through coding challenges, answer
questions, and try to foster a deeper understanding of programming concepts.

The study conducted by \cite{chan23} revealed that both undergraduate and
postgraduate students have positive attitudes towards the use of GenAI.
A systematic review organized by \cite{Lo24} demonstrated that students can
effectively learn from ChatGPT, resulting in improved comprehension and academic
achievement \citep{Callejo24}. Furthermore, it was observed that ChatGPT may
allow students to regulate their learning pace and can potentially support
self-regulated learning, particularly for students with prior technical and
disciplinary knowledge \citep{Xia23}.

@@

@@

During the categorization phase, the students' prompts were initially classified
into eight distinct categories using the Claude.ai \citep{claude} tool.
Subsequently, the authors of this study manually reviewed these classifications
to ensure accuracy and alignment with the research objectives.

\begin{table*}[htbp]
    \caption{Categories - adapted from \cite{Ghimire24}}
    \begin{center}
      \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
      \begin{tabular}{p{3cm} p{4cm} p{3cm}} % Remove vertical bars
        \hline
        \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
        \hline
        Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
        Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
        Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
        Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
        Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
        Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
        Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
        Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
        \hline
      \end{tabular}
      \label{tab:categories}
    \end{center}
  \end{table*}

Initially, \cite{Ghimire24} proposed four categories: Debugging Help (DH),
Conceptual Question (CQ), Code Snippet (CS), and Complete Solution (CSO).
However, the data collected indicated the need for additional categories,
leading to the inclusion of four more: Multiple Questions (MQ), Student
Corrections (SC), Language Change (LC), and Uncategorized (U). Unlike study
\cite{Ghimire24}, which analyzed students' behavior using artificial
intelligence for specific programming tasks, our study allowed participants to
explore the chatbot as they preferred. This approach enabled the emergence of
additional categories of analysis, broadening the understanding of the tool's
uses and interactions. Table \ref{tab:categories} presents these categories,
their descriptions, and an example. Besides that, the topics covered in the
students' prompts were categorized based on the course issues, which include
(1) Java, (2) Object Orientation, (3) Testing, (4) Recursion, (5) Data
Structures, and (6) Sorting and Searching Algorithms.

>>>>

Recent advances in Generative Artificial Intelligence (GenAI) have created new
opportunities in professional programming and education \citep{Puryear22}. In
programming courses, students can employ GenAI tools to improve their
understanding, receive personalized feedback, and access detailed explanations.
For example, GitHub Copilot, a tool integrated into development environments,
assists in providing real-time suggestions and accelerating code writing, which
could help streamline the learning process. On the other hand, educational
chatbots try to guide students through coding challenges, answer questions, and
foster a deeper understanding of programming concepts.

The study by \cite{chan23} revealed that undergraduate and postgraduate students
have positive attitudes towards using GenAI. A systematic review organized by
\cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
resulting in improved comprehension and academic achievement \citep{Callejo24}.
Additionally, it was observed that ChatGPT can enable students to control their
learning pace and may support self-regulated learning, especially for those with
prior technical and disciplinary knowledge \citep{Xia23}.

@@

\begin{table*}[htbp]
  \caption{Categories - adapted from \cite{Ghimire24}}
  \begin{center}
    \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
    \begin{tabular}{p{4.1cm} p{4cm} p{3.7cm}} % Remove vertical bars
      \hline
      \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
      \hline
      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
      Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
      Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
      Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
      \hline
    \end{tabular}
    \label{tab:categories}
  \end{center}
\end{table*}

@@

During the categorization phase, the students' prompts were initially classified
into eight distinct categories using the Claude.ai \citep{claude} tool.
Subsequently, the authors of this study manually reviewed these classifications
to ensure accuracy and alignment with the research objectives.
Initially, \cite{Ghimire24} proposed four categories: Debugging Help (DH),
Conceptual Question (CQ), Code Snippet (CS), and Complete Solution (CSO).
However, the data collected indicated the need for additional categories,
leading to the inclusion of four more: Multiple Questions (MQ), Student
Corrections (SC), Language Change (LC), and Uncategorized (U). Unlike study
\cite{Ghimire24}, which analyzed students' behavior using artificial
intelligence for specific programming tasks, our study allowed participants to
explore the chatbot as they preferred. This approach enabled the emergence of
additional categories of analysis, broadening the understanding of the tool's
uses and interactions. Table \ref{tab:categories} presents these categories,
their descriptions, and an example. Besides that, the topics covered in the
students' prompts were categorized based on the course issues, which include
(1) Java, (2) Object Orientation, (3) Testing, (4) Recursion, (5) Data
Structures, and (6) Sorting and Searching Algorithms.
